---
title: "p8105_hw2_sl4655"
author: "Shuya Liu"
date: "October 2, 2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
```

# _Problem 1_

### Read and clean the Mr. Trash Wheel sheet

```{r trash_wheel_data_input, message = FALSE}
## Read in Mr. Trash Wheel data
trash_wheel_data <- 
  readxl::read_excel(path = "./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
                     sheet = "Mr. Trash Wheel") %>%
  janitor::clean_names() %>%
  select(dumpster:homes_powered) %>%
  drop_na(dumpster) %>%
  mutate(sports_balls = round(sports_balls) %>% as.integer())
```

### Read and clean precipitation data for 2017 and 2018

```{r precipitaion}
precip_2017_data <- 
  readxl::read_excel(path = "./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
                     sheet = "2017 Precipitation",
                     range = "A2:B14") %>%
  janitor::clean_names() %>%
  drop_na(total) %>%
  arrange(month) %>%
  mutate(year = 2017, month = month.name)

precip_2018_data <-
  readxl::read_excel(path = "./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
                     sheet = "2018 Precipitation",
                     range = "A2:B14") %>%
  janitor::clean_names() %>%
  drop_na(total) %>%
  arrange(month) %>%
  mutate(year = 2018, month = month.name)
```

### Combine precipitation datasets

```{r combine}
precip_combine_data <-
  rbind(precip_2017_data, precip_2018_data) %>%
  select(year, month, precipitation = total)
```

```{r stats, include = FALSE}
## Descriptive Statistics
total_2017 = sum(with(precip_combine_data, subset(precipitation, year == 2017)))
total_2018 = sum(with(precip_combine_data, subset(precipitation, year == 2018)))
min_2017 = min(with(precip_combine_data, subset(precipitation, year == 2017)))
min_2018 = min(with(precip_combine_data, subset(precipitation, year == 2018)))
max_2017 = max(with(precip_combine_data, subset(precipitation, year == 2017)))
max_2018 = max(with(precip_combine_data, subset(precipitation, year == 2018)))
```

* **For precipitation data**:

  The total precipitation for year 2017 is `r total_2017`(in). The total precipitation for year 2018 is `r total_2018`(in). The difference is `r total_2018 - total_2017`(in). In year 2017, the smallest amount of precipitation `r min_2017`(in) occurred in `r with(precip_combine_data, subset(month, precipitation == min_2017))`, and the largest amount of precipitation `r max_2017`(in) occurred in `r with(precip_combine_data, subset(month, precipitation == max_2017))`. In year 2018, the smallest amount of precipitation `r min_2018`(in) occurred in `r with(precip_combine_data, subset(month, precipitation == min_2018))`, and the largest amount of precipitation `r max_2018`(in) occurred in `r with(precip_combine_data, subset(month, precipitation == max_2018))`.

* **For Mr. Trash Wheel data**:

  There are `r dim(trash_wheel_data)[1]` rows of data and `r dim(trash_wheel_data)[2]` variables, with `r sum(is.na(trash_wheel_data))` missing data. The data starts from `r min(trash_wheel_data$date)` to `r max(trash_wheel_data$date)`. The medians of plastic bottles, cigarette butts, grocery bags and sports balls for each year are shown as below:
  
```{r echo = FALSE}
aggregate(
  cbind(plastic_bottles, cigarette_butts, grocery_bags, sports_balls) ~ year, 
  data = trash_wheel_data, 
  median)
```

The total amount of weight(tons), and volume(cubic_yards), and homes powered for each year are shown as below:

```{r echo = FALSE}
aggregate(
  cbind(weight_tons, volume_cubic_yards, homes_powered) ~ year,
  data = trash_wheel_data,
  sum
)
```

# _Problem 2_

### Clean the data in pols-month.csv
```{r pols_month}
pols_data <- 
  read_csv(file = "./data/fivethirtyeight_datasets/pols-month.csv") %>%
  janitor::clean_names()
```



First, clean the data in pols-month.csv. Use separate() to break up the variable mon into integer variables year, month, and day; replace month number with month name; create a president variable taking values gop and dem, and remove prez_dem and prez_gop; and remove the day variable.

Second, clean the data in snp.csv using a similar process to the above. For consistency across datasets, arrange according to year and month, and organize so that year and month are the leading columns.

Third, tidy the unemployment data so that it can be merged with the previous datasets. This process will involve switching from “wide” to “long” format; ensuring that key variables have the same name; and ensuring that key variables take the same values.

Join the datasets by merging snp into pols, and merging unemployment into the result.

Write a short paragraph about these datasets. Explain briefly what each dataset contained, and describe the resulting dataset (e.g. give the dimension, range of years, and names of key variables).

Note: we could have used a date variable as a key instead of creating year and month keys; doing so would help with some kinds of plotting, and be a more accurate representation of the data. Date formats are tricky, though. For more information check out the lubridate package in the tidyverse.